Assessing the impact of urban heating on tree growth in Berlin with open inventory and environmental data

### 2022-01-12 analysis/paper/paper.Rmd
disc

- Quercus robur, Tilia intermedia and Aesculum hippocastanum performed best, matching with climate-species matrix in roloff and brune		
- least: Platanus acerifolia, Acer pseudoplatanus.. data issues or location specific factors where trees were planted. Avenue for further work to confirm and identify drivers.
- Urban fabric: LCZ density 
- and sensitive trees in hot spots

@@

**Increased growth rates.. live fast, die young - high turnover! requires focused planting and/or maintenance to counter [@smith2019].**

>>>>

Generally, the ten investigated species and cultivars showed consistent responses in sign, but differed substantially in magnitude, modulated by age.
*Aesculum hippocastanum* , *Tilia intermedia*, *T. intermedia 'pallida'* and showed greatest absolute growth at high temperatures at younger (30-35) as well as older ages (60-65).
The drought tolerance and performance under high temperatures is well documented for the latter two and features in tree selection frameworks, such as species climate matrices in @roloff2009 and in @brune2016, where several studies investigating (hydro)climatic sensitivity were summarized.
Contrastingly, *Acer pseudoplatanus* showed decreased absolute growth with higher temperatures; this corresponds with their classification as *not very suitable* under [@roloff2009].
*Platanus acerifolia* and *Acer platanoides* are considered *suitable* and *very suitable* under @roloff2009, yet show the greatest decline with temperatures here.
Indeed, several several studies reviewed in @brune2016 highlight *A. platanoides* as drought-tolerant.
Such contrasting indications are also reported for *A. pseudoplatanus*, and even report *A. hippocastanum* as only moderately tolerant, despite performing best here. 
These differences may be due to idiosyncrasies of planting sites, such as low soil volume and increased potential for water stress, (long-term) interactions with pathogens and herbivores [e.g., @meineke2018], or perhaps related to data quality.
Most important, these differences highlight the need for assessments of performance (e.g., as growth) under location-specific conditions, and this study provides a first attempt and insights into temperature sensitivity at the scale of an entire metropolis.

- Urban fabric: LCZ density 
- and sensitive trees in hot spots ..**Increased growth rates.. live fast, die young - high turnover! requires focused planting and/or maintenance to counter [@smith2019].**

@@
